---
title: "BIOS 611 HOMEWORK 9"
author: "Ally Wardell"
date: "10/19/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse')
library('stats')
library('matlab')
library('pROC')
library('MASS')
library('gbm')
```

## DATA
Grab the powers data sets from:

https://github.com/Vincent-Toups/datasci611/raw/main/lectures/12-classification/source_data/prime_earth_characters.csv https://github.com/Vincent-Toups/datasci611/raw/main/lectures/12-classification/source_data/prime_earth_powers.csv

## QUESTION 1
Reproduce the code which replaces unusual powers with "other", and pivots the resulting data set to a wide form. You can copy the code from lecture but this should be something you can do without doing so now.

```{r make dataset}
genders <- read_csv("https://raw.githubusercontent.com/Vincent-Toups/datasci611/main/lectures/12-classification/source_data/prime_earth_characters.csv") %>%
    filter(property_name=="gender") %>% select(-property_name, -universe) %>%
    rename(gender=value);

powers <- read_csv("https://raw.githubusercontent.com/Vincent-Toups/datasci611/main/lectures/12-classification/source_data/prime_earth_powers.csv") %>%
    select(power, character);

power_counts <- powers %>% group_by(power) %>% tally() %>% arrange(desc(n));
common <- power_counts %>% pull(power) %>% head(20);
uncommon <- power_counts %>% filter(!(power %in% common)) %>% pull(power);

powers_wide <- powers %>% mutate(power = {
    p <- power;
    p[p %in% uncommon] <- "other";
    p
}) %>% distinct() %>%
    mutate(dummy=1) %>%
    pivot_wider(id_cols="character", names_from="power", values_from="dummy", values_fill=list(dummy=0));

data <- genders %>%
    inner_join(powers_wide, by="character") %>%
    filter(gender %in% c("male", "female")) %>%
    mutate(is_female=1*(gender=="female"))

```

## QUESTION 1A
a. Use a dimensionality reduction method of your choice to project the data down to six dimensions.
```{r dimensionality reduction}

powers_wide_1a <- data %>% select(-character, -gender, -is_female) 

pca_results <- prcomp(powers_wide_1a)
pca_results
imagesc(pca_results$x)
summary(pca_results)

pca_results_df <- data.frame(pca_results$x) %>% select(PC1, PC2, PC3, PC4, PC5, PC6)

pca_df <- cbind(pca_results_df, data %>% select(is_female))

```

## QUESTION 1B
b. use LDA or Logistic Regression to produce a clustering.
``` {r LDA}
select <- dplyr::select;

lda_results <- lda(is_female~., data=pca_df);

probability <- predict(lda_results, newdata = pca_df, type="prob")$posterior

probability_2 <- as_tibble(probability) %>% select(2) %>% rename(is_female_prob=1)

powers_wide_1b <-cbind(pca_df, probability_2)

category_test_df <-powers_wide_1b %>% 
  mutate(is_female_pred = ifelse(is_female_prob>0.5, 1, 0))

category_test_df %>% group_by(is_female, is_female_pred) %>% tally()

category_test_df

```

## QUESTION 1C
Calculate and plot an ROC curve. (Note that arrange I'm using to make the ROC curve look nice).

```{r q1c}

roc(category_test_df$is_female ~ category_test_df$is_female_prob, plot = TRUE, print.auc = TRUE, legacy.axes=TRUE)

```

##QUESTION 2
Reproduce the GBM classifier from lecture. Using your test data set calculate the

    i. true positive, false positive, true negative, false negative rates
    ii. the precision, recall and f1 score.
    iii. plot an ROC Curve.
    
```{r q2}
explanatory <- data %>% select(-character, -gender, -is_female) %>% names()

formula <- as.formula(sprintf("is_female ~ %s", paste(explanatory, collapse=" + ")))

tts <- runif(nrow(data)) < 0.5

train <- data %>% filter(tts)

test <- data %>% filter(!tts)


model <- gbm(formula, data=train)
prob <- predict(model, newdata=test, type="response")
test_q2 <- test %>% mutate(is_female_pred=1*(prob>0.5))
test_q2 %>% group_by(is_female, is_female_pred) %>% tally()

#i. CALCULATE RATES
#1. true positive rate: how often are positive cases classified as positive?
#2. true negative rate: how often are negative cases classified as negative?
#3. false positive rate: how often are negative cases classified as positive?
#4. false negative rate: how often are positive cases classified as negative?
#5  accuracy: The proportion classified correctly.
#6. recall or sensitivity: the number of true positives divided by the
#   total number of positives. You can also calculate the true negative
#   rate similarly.
#7. precision: true positive count over the number of all objects classified as positive.

rates <- function(actual, predicted) { 
  positive <- which(!!actual)
  negative <- which(!actual)
  
  true_positive_rate <- sum(predicted[positive])/length(positive)
  false_positive_rate <- sum(predicted[negative])/length(negative)
  
   true_negative_rate <- sum(!predicted[negative])/length(negative)
   false_negative_rate <- sum(!predicted[positive])/length(positive)

accuracy <- (true_positive_rate + true_negative_rate)/(true_positive_rate + true_negative_rate + false_positive_rate + false_negative_rate)

recall <- ((true_positive_rate)/ (true_postitive_rate + false_negative_rate))
precision <- ((true_positive_rate)/(true_positive_rate + false_positive_rate))

f1_score <- 2*((precision*recall)/(precision + recall))

  tibble(true_positive_rate = true_positive_rate,
         false_positive_rate = false_positive_rate,
         true_negative_rate = true_negative_rate,
         false_negative_rate = false_negative_rate,
         accuracy = sum(actual==predicted)/length(actual),
         accuracy2 = accuracy,
         recall = recall,
         precision = precision,
         f1_score = f1_score)

}
rates(test_q2$is_female, test_q2$is_female_pred)


#iii plot ROC curve

roc(test_q2$is_female ~ test_q2$is_female_pred, plot = TRUE, print.auc = TRUE, legacy.axes = TRUE)


# # ROC ANOTHER WAY
# rate <- function(a){
#     sum(a)/length(a);
# }
# maprbind <- function(f,l){
#     do.call(rbind, Map(f, l));
# }
# 
# roc <- maprbind(function(thresh){
#     ltest <- test_ex %>% mutate(female_pred=1*(is_female_pred>=thresh)) %>%
#         mutate(correct=female_pred == is_female);
#     tp <- ltest %>% filter(ltest$is_female==1) %>% pull(correct) %>% rate();
#     fp <- ltest %>% filter(ltest$is_female==0) %>% pull(correct) %>% `!`() %>% rate();
#     tibble(threshold=thresh, true_positive=tp, false_positive=fp);
# }, seq(from=0, to=1, length.out=10)) %>% arrange(false_positive, true_positive)
# ggplot(roc, aes(false_positive, true_positive)) + geom_line();

```

##QUESTION 3
Now consider traditional gender vs non-traditional gender. Treat "male" and "female" as "traditional" and all other genders as "non-traditional". Build a gbm to classify gender as traditional vs non-traditional using super powers.

HINT: when you perform you train test split, make sure you do so in such a way that both the training and testing set contain examples of the relatively rare non-traditional genders.

Calculate:

true positive, false positive, true negative, false negative rates
the precision, recall and f1 score.
plot an ROC Curve.

``` {r Q3}

data2 <- genders %>%
    inner_join(powers_wide, by="character") %>%
    mutate(across(everything(), factor)) %>% 
    mutate(gender_trad = if_else(gender=="male" | gender == "female", 1, 0))

explanatory <- data2 %>% select(-character, -gender, -gender_trad) %>% names()
formula3 <- as.formula(sprintf("gender_trad ~ %s", paste(explanatory, collapse=" + ")));

data_trad <- data2 %>% filter(gender_trad==1)

data_non_trad <- data2 %>% filter(gender_trad==0)

tts_trad <- runif(nrow(data_trad)) < 0.5;
tts_non_trad <- runif(nrow(data_non_trad )) < 0.5;

train_trad <- data_trad %>% filter(tts_trad) 
train_non_trad <- data_non_trad %>% filter(tts_non_trad) 
test_trad <- data_trad %>% filter(!tts_trad) 
test_non_trad <- data_non_trad %>% filter(!tts_non_trad) 

train3 <- rbind(train_trad, train_non_trad)
test3 <- rbind(test_trad, test_non_trad)

model <- gbm(formula3, data=train3);
prob <- predict(model, newdata=test3, type="response");
test_q3 <- test3 %>% mutate(gender_trad_pred=1*(prob>0.5));
test_q3 %>% group_by(gender_trad, gender_trad_pred) %>% tally()


#Calculate:

#true positive, false positive, true negative, false negative rates
#the precision, recall and f1 score.
rates(test_q3$gender_trad, test_q3$gender_trad_pred)

roc(test_q3$gender_trad ~ test_q3$gender_trad_pred, plot = TRUE, print.auc = TRUE, legacy.axes = TRUE)


```